---
title: "Lab 09 - Grading the professor, Pt. 1"
author: "Elayna Seago"
date: "03/15/22"
output: github_document
---

### Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
library(tidymodels)
library(openintro)
```

### Exercise 1
Yes, the distribution of scores is skewed negatively. The majority of students evaluate professors as performing well. I expected to see more of a binomial distribution because I thought students who really liked or really dislike the course would be the primary respondents.
```{r}
evals %>%
  ggplot(aes(x = score)) +
  geom_histogram()
```


### Exercise 2
There does not really appear to be much of a linear relationship here.
```{r}
evals %>%
  ggplot(aes(x = bty_avg , y = score)) +
  geom_point()
```

# Exercise 3
The original graph hid some repeat points on the right side of the graph. The relationship looks a little more positive and linear now. Jitter adds a little random shift to the points to minimize overlap.
```{r}
evals %>%
  ggplot(aes(x = bty_avg , y = score)) +
  geom_point()+
  geom_jitter()
```

# Exercise 4
score = 3.88 + bty_avg*.067
```{r}
m_bty <- lm(score ~ bty_avg, data=evals)  

print(m_bty)

summary(m_bty)
```

# Exercise 5

```{r}
evals %>%
  ggplot(aes(x = bty_avg , y = score)) +
  geom_point()+
  geom_jitter()+
  geom_smooth( method = lm , se = FALSE , color = "orange")
```

# Exercise 6
The slope is .0664, so for every 1 point a professors attractiveness rating goes up, there overall evaluation score is expected to increase .0664.
# Exercise 7
The intercept is 3.88, so a professor with an attractiveness rating of 0 would still be expected to have an overall rating of 3.88.
# Exercise 8

```{r}
summary(m_bty)$r.squared
```


The R squared is .035 and it tells us how much variation in overall score is explained by beauty rating.

# Exercise 9

```{r gender-model}
m_gen <- lm(score ~ gender, data=evals)  

print(m_gen)

summary(m_gen)
```

# Exercise 10

males:
score = 4.09 + (1*.1415)

females:
score = 4.09 + (0*.1415)

```{r gender-plot}
evals %>%
  ggplot(aes(x = gender , y = score ,  color = gender)) +
  geom_point()
```


# Exercise 11

#score = 4.28 + (-.1297*x1) + (-.1452*x2)

teaching track:
score= 4.28 + (0) + (0)

tenure track:
score = 4.28 + (1*-.1297) + (0)

tenured:
score = 4.28 + (0) + (1*-.1452)

the intercept = 4.28, this is a professors score at level 0 0 (aka teaching track)
slope 1 = -.1297, this is how much a professors score decreases at level 1 0 (in the contrast matrix, this is a tenure track professor)
slope 2 = -.1452, this is how much a professors score decreases at level 0 1 (tenured)

```{r rank-model}
m_rank <- lm(score ~ rank, data=evals)  

print(m_rank)

summary(m_rank)
```

```{r}
head(model.matrix(m_rank), 20) 
```


```{r plot-for-confirmation}
evals %>%
  ggplot(aes(x = rank , y = score ,  color = gender)) +
  geom_point()+
  geom_jitter()
```



# Exercise 12

```{r relevel}
evals_relevel <- evals %>% 
  mutate(rank_relevel = relevel(rank, ref = "tenure track"))
```


# Exercise 13

```{r relevel-model}

m_rank_relevel <- lm(score ~ rank_relevel, data=evals_relevel)
summary(m_rank_relevel)

```

#initial lm
teaching track:
score= 4.28 + (0)=4.28

tenure track:
score = 4.28 + (1*-.1297)=4.15

tenured:
score = 4.28 + (1*-.1452)=4.13

#new lm
teaching track:
score= 4.15 + (1*.129)=4.279

tenure track:
score= 4.15 + (0)=4.15

tenured:
score= 4.15 + (1*-.015)=4.13

the intercept = 4.15, this is a professors score at level 0 0 (aka tenure track)
slope 1 = .1297, this is how much a professors score increases at level 1 0 (in the new contrast matrix, this is a teaching track professor)
slope 2 = -.015, this is how much a professors score decreases at level 0 1 (tenured)
R2=.007, this means the linear model shows that teaching position explains .7% of the variance in score.

# exercise 14

```{r recode}
evals <- evals %>% 
  mutate(tenure_eligible = recode(rank, "tenure track" = "yes" ,
                                  "tenured" = "yes" ,
                                  "teaching" = "no"
  ))
```

```{r confirm-it-worked}
head(evals)
```

#exercise 15

```{r}
m_tenure_eligible <- lm(score ~ tenure_eligible, data=evals)
summary(m_tenure_eligible)
```


score = 4.28 + (-.14*x)
for tenure eligible, that means 4.14 and for tenure non-eligible, 4.28. The slope is the amount that score goes down for those who are tenure eligible. Since tenure eligibility is categorical, this shows the average score for those two groups.
R-squared is .009, so tenure eligibility explains .9% of the variance in score.

